On the Functionality of the Multilayer Perceptron

### 2020-12-17 Thesis/Drafts/Buell-PHYS799-ThesisDraft3.tex
Added Draft 4, Fixed some figures

\bibitem{UnivIowa}
University of Iowa Electronic Music Studios home page- \textit{http://theremin.music.uiowa.edu/}

\bibitem{White}
White, Harvey Elliott, and Donald H. White. \textit{Physics and Music: the Science of Musical Sound}. Dover Publications, Inc., 2019.

\bibitem{Zhang}
Zhang, Tong, and C.-C. Jay Kuo. “Content-Based Classification and Retrieval of Audio.” \textit{Advanced Signal Processing Algorithms, Architectures, and Implementations VIII}, 2 Oct. 1998, pp. 432–443., doi:10.1117/12.325703.

>>>>

^{•}

### 2020-12-17 Thesis/Drafts/Buell-PHYS799-ThesisDraft4.tex
Added Draft 4, Fixed some figures

\paragraph*{}We execute a K-Folds cross validation algorithm to evaluate the performance of the neural network across overlapping subsets of data. This ensures that the model can perform consistently and reliably given small changes in initial conditions. In this stage, we also examine the value of hyper-parameters, activation functions, and layer widths to best compliment the properties of the data set. This process is repeated and expanded upon until we have produced a model with a sufficient performance. We explore this in section (\ref{sec-PerfEval}). Results of all metrics are computed with the \textit{numpy},\textit{scipy}, and \textit{scikit learn} python modules \cite{numpy,scipy,sklearn}. All performance visualizations are produced with the \textit{matplotlib} python package \cite{matplotlib}.

@@

\paragraph*{}A Neural Network is a model of a mathematical function, composed of several smaller mathematical functions called \textit{layers} \cite{Goodfellow,Loy}. Each layer represents an operation that takes an input, and returns an output . The exact nature of this operation can be very different depending on the layer type or where it sits within the network. Typically, a neural network is characterized by a chain-like structure where the output of one layer is fed directly into the input of the next layer \cite{Geron,Goodfellow}. It is this process of transforming inputs successively in a particular order until an output is attained that makes up the core functionality of a neural network \cite{Geron,Loy}. The output of the final layer encodes the "decision" of the model given a particular input.

\paragraph*{}Other than inspiration from the brain, Neural Networks are named \textit{networks} due to their successive chained or nested composition nature. A standard representation of a neural network is often a linked list or computational graph structure \cite{Goodfellow}. This maps out exactly how the repeated compositions are structured and organizes the flow of information in more complicated networks. Each node in the computational graph represents a \textit{layer}, which executes a particular prescribed mathematical function \cite{Goodfellow}. In the case of a \textit{linear, feed-forward} network, such as the one that we implement, information passes in a single direction to produce an output. We can represent this as a graph:
\begin{equation}
\label{eqn-FunctionGraph}
F^*(x) = x \rightarrow f^{(0)} \rightarrow f^{(1)} \rightarrow ..... \rightarrow
f^{(L-2)} \rightarrow f^{(L-1)} \rightarrow y^*
\end{equation}
or as a nested function:
\begin{equation}
\label{eqn-FunctionChain}
F^*(x) = f^{(L-1)}\big[f^{(L-2)}\big[...f^{(1)}\big[f^{(0)}\big[x\big]\big]\big]...\big]\big] = y^*
\end{equation}
The number of layers in a neural network is referred to as the network \textit{depth}. The dimensionality of each layer is referred to as the network \textit{width} \cite{Geron,Loy}.
A network model that contains $L$ unique layers is said to be an $L$-Layer Neural Network, with each layer indexed by a superscipt, $(0)$ through $(L-1)$. Layer $(0)$ is said to be the \textit{input layer} and layer $(L-1)$ is said to be the \textit{output layer} \cite{Geron,Loy}.

@@

\paragraph*{}As stated previously, a neural network is composed of a series of functions that are called in succession to transform features (an input) into predictions (an output). As shown in Eq. (\ref{eqn-altLayerFunction}), each function feeds directly into the next as to form a chain-like computational graph \cite{Goodfellow}. There are multiple type of layers that we use to construct the network in this project, but the two most important are called a \textit{dense} layer and a \textit{convoltuion} layer. These layers each contain a set of parameters which makeupa  subset of the trainable elements in $\Theta$.
These two layer functions can be divided into two portions: (i.) a Linear transformation, with a bias addition, and (ii.) an element-wise activation transformation. The Linear transformation can be given by Eq.(\ref{eqn-LinearTransform}) in the case of a 1D Dense layer (\ref{subsubsec-DenseLayer}) or Eq. (\ref{eqn-LinearTransform2}) in the case of a 2D Convolution layer.
\begin{equation}
\label{eqn-LinearTransform}
z^{(l)} = W^{(l)} x^{(l-1)} + b^{(l)}
\end{equation}
\begin{equation}
\label{eqn-LinearTransform2}
z^{(l)} = \big(W^{(l)} * x^{(l-1)}\big) + b^{(l)}
\end{equation} 
Where $W^{(l)} x^{(l-1)}$ denotes a matrix-vector product, and $\big(W^{(l)} * x^{(l-1)}\big)$ denotes a convolution product. In both cases, $W^{(l)}$ is a \textit{weighting-matrix} or \textit{weighting-kernel} for layer $(l)$, $b^{(l)}$ is the \textit{bias-vector} for layer $(l)$, $z^{(l)}$ are the \textit{linear-activations} for layer $(l)$ and $x^{(l-1)}$ is the final activations for layer $(l-1)$.

@@

\paragraph*{}The wave equation is a \textit{linear} partial differential equation which allows for a linear combination of solutions to form a sound wave. In the case of the taught finite string that is fixed at both ends, such a guitar, viola or cello, the a solution tends to take the form of a sine or cosine function \cite{Hunter, Taylor, White}. Notice the terms arise by the many a parallel yellow and orange lines in each spectrogram in Fig. (\ref{fig-PropertiesChordophone}). The verical position of any line is given by the linear frequency of the sinusoid, and the brightness is proportional to the coefficient of the combination. The boundary conditions, and most forms of standard initial conditions allow for the each sinusoidal to fit exactly an integer or half-integer number wavelengths into the region $0 \leq x \leq L$ \cite{Haberman,Hunter}. Players of chordophones change pitch by shortening the length of the vibrating string or applying additional tension to the string. This changes the length of portion of the string that vibrates, and causes the production of a new pitch.

@@

\paragraph*{}Aerophones all produce sound by exciting molecules of air within a column that vibrates in different patterns called \textit{modes} \cite{White}. the wavelength and energy of each mode is  determined by the structure of the instruments, and the actions of the player. Most aerophones such a flutes, clarinets, and saxophones are more much more reminiscent of air columns that are closed at one end and open at the other. This is marked by the presence of a mouth piece at one end where a played blows air from their lungs to add energy, and a bell at the other end to serve as an amplification structure \cite{Olson}. For a column of length $L$, that is open at one end, the fundamental frequency of the vibration is given by \cite{Olson}:
\begin{equation}
\label{eqn-OpenClosedCol}
f_0 = \frac{v}{4L} = \frac{v}{\lambda}
\end{equation}
where $v$ is the velocity of sound in the medium, and $\lambda$ is the wavelength of the sound wave. Eq. (\cite{eqn-OpenClosedCol}) gives the lowest pitch that such an instrument with a length $L$ can play. Lower woodwinds such as bass clarinets, have a much longer vibrating air column that that of the much shorter piccolo, thus can produce considerably lower notes.

@@

^{•}

>>>>

\paragraph*{}We execute a K-Folds cross validation algorithm to evaluate the performance of the neural network across overlapping subsets of data. This ensures that the model can perform consistently and reliably given small changes in initial conditions. In this stage, we also examine the value of hyper-parameters, activation functions, and layer widths to best compliment the properties of the data set. This process is repeated and expanded upon until we have produced a model with a sufficient performance. We explore this in section (\ref{sec-PerfEval}). Results of all metrics are computed with the \textit{numpy},\textit{scipy}, and \textit{scikit learn} python modules \cite{Numpy,Scipy,Sklearn}. All performance visualizations are produced with the \textit{matplotlib} python package \cite{Matplotlib}.

@@

\paragraph*{}A neural network is a model of a mathematical function, composed of several smaller mathematical functions called \textit{layers} \cite{Goodfellow,Loy}. Each layer represents an operation that takes an input, and returns an output . The exact nature of this operation can be very different depending on the layer type or where it sits within the network. Typically, a neural network is characterized by a chain-like structure where the output of one layer is fed directly into the input of the next layer \cite{Geron,Goodfellow}. It is this process of transforming inputs successively in a particular order until an output is attained that makes up the core functionality of a neural network \cite{Geron,Loy}. The output of the final layer encodes the "decision" of the model given a particular input.

\paragraph*{}Other than inspiration from the brain, neural networks are named \textit{networks} due to their successive chained or nested composition nature. A standard representation is often based on a linked list or computational graph structure \cite{Goodfellow}. This maps out exactly how the repeated compositions are structured and organizes the flow of information in more complicated networks. Each node in the computational graph represents a \textit{layer}, which executes a particular prescribed mathematical function \cite{Goodfellow}. In the case of a \textit{linear, feed-forward} network, such as the one that we implement, information passes in a single direction to produce an output. We can represent this as a graph:
\begin{equation}
\label{eqn-FunctionGraph}
F^*(x) = x \rightarrow f^{(0)} \rightarrow f^{(1)} \rightarrow ..... \rightarrow
f^{(L-2)} \rightarrow f^{(L-1)} \rightarrow y^*
\end{equation}
or as a nested function:
\begin{equation}
\label{eqn-FunctionChain}
F^*(x) = f^{(L-1)}\big[f^{(L-2)}\big[...f^{(1)}\big[f^{(0)}\big[x\big]\big]\big]...\big]\big] = y^*
\end{equation}
The number of layers in a neural network is referred to as the network \textit{depth}. The dimensionality of each layer is referred to as the network \textit{width} \cite{Geron,Loy}.
A network model that contains $L$ unique layers is said to be an $L$-Layer Neural Network, with each layer indexed by a superscipt, $(0)$ through $(L-1)$. Layer $(0)$ is said to be the \textit{input layer} and layer $(L-1)$ is said to be the \textit{output layer} \cite{Geron,Loy}.

@@

\paragraph*{}As stated previously, a neural network is composed of a series of functions that are called in succession to transform features (an input) into predictions (an output). As shown in Eq. (\ref{eqn-altLayerFunction}), each function feeds directly into the next as to form a chain-like computational graph \cite{Goodfellow}. There are multiple type of layers that we use to construct the network in this project, but the two most important are called a \textit{dense} layer and a \textit{convolution} layer. These layers each contain a set of parameters which makeup a subset of the trainable elements in $\Theta$.
These two layer functions can be divided into two portions: (i.) a Linear transformation, with a bias addition, and (ii.) an element-wise activation transformation. The Linear transformation can be given by Eq.(\ref{eqn-LinearTransform}) in the case of a 1D Dense layer (\ref{subsubsec-DenseLayer}) or Eq. (\ref{eqn-LinearTransform2}) in the case of a 2D Convolution layer.
\begin{equation}
\label{eqn-LinearTransform}
z^{(l)} = W^{(l)} x^{(l-1)} + b^{(l)}
\end{equation}
\begin{equation}
\label{eqn-LinearTransform2}
z^{(l)} = \big(W^{(l)} * x^{(l-1)}\big) + b^{(l)}
\end{equation} 
Where $W^{(l)} x^{(l-1)}$ denotes a matrix-vector product, and $\big(W^{(l)} * x^{(l-1)}\big)$ denotes a convolution product. In both cases, $W^{(l)}$ is a \textit{weighting-matrix} or \textit{weighting-kernel} for layer $(l)$, $b^{(l)}$ is the \textit{bias-vector} for layer $(l)$, $z^{(l)}$ are the \textit{linear-activations} for layer $(l)$ and $x^{(l-1)}$ is the final activations for layer $(l-1)$.

@@

\paragraph*{}The wave equation is a \textit{linear} partial differential equation which allows for a linear combination of solutions to form a sound wave. In the case of the taught finite string that is fixed at both ends, such a guitar, viola or cello, the a solution tends to take the form of a sine or cosine function \cite{Hunter, Taylor, White}. Notice the terms arise by the many a parallel yellow and orange lines in each spectrogram in Fig. (\ref{fig-PropertiesChordophones}). The verical position of any line is given by the linear frequency of the sinusoid, and the brightness is proportional to the coefficient of the combination. The boundary conditions, and most forms of standard initial conditions allow for the each sinusoidal to fit exactly an integer or half-integer number wavelengths into the region $0 \leq x \leq L$ \cite{Haberman,Hunter}. Players of chordophones change pitch by shortening the length of the vibrating string or applying additional tension to the string. This changes the length of portion of the string that vibrates, and causes the production of a new pitch.

@@

\paragraph*{}Aerophones all produce sound by exciting molecules of air within a column that vibrates in different patterns called \textit{modes} \cite{White}. the wavelength and energy of each mode is  determined by the structure of the instruments, and the actions of the player. Most aerophones such a flutes, clarinets, and saxophones are more much more reminiscent of air columns that are closed at one end and open at the other. This is marked by the presence of a mouth piece at one end where a played blows air from their lungs to add energy, and a bell at the other end to serve as an amplification structure \cite{Olson}. For a column of length $L$, that is open at one end, the fundamental frequency of the vibration is given by \cite{Olson}:
\begin{equation}
\label{eqn-OpenClosedCol}
f_0 = \frac{v}{4L} = \frac{v}{\lambda}
\end{equation}
where $v$ is the velocity of sound in the medium, and $\lambda$ is the wavelength of the sound wave. Eq. (\ref{eqn-OpenClosedCol}) gives the lowest pitch that such an instrument with a length $L$ can play. Lower woodwinds such as bass clarinets, have a much longer vibrating air column that that of the much shorter piccolo, thus can produce considerably lower notes.

@@

\bibitem{White}
White, Harvey Elliott, and Donald H. White. \textit{Physics and Music: the Science of Musical Sound}. Dover Publications, Inc., 2019.

\bibitem{UnivIowa}
University of Iowa Electronic Music Studios home page- \textit{http://theremin.music.uiowa.edu/}

\bibitem{Zhang}
Zhang, Tong, and C.-C. Jay Kuo. “Content-Based Classification and Retrieval of Audio.” \textit{Advanced Signal Processing Algorithms, Architectures, and Implementations VIII}, 2 Oct. 1998, pp. 432–443., doi:10.1117/12.325703.
